On the Dynamics of Anthropomorphism in Robotics

### 2013-07-08 dynamics-anthropomorphism.tex
Rewrite/complete section on familiarization

\paragraph{Disruptive behaviors}
\ref{sec:disruptive}

>>>>

\emph{Familiarization}, in this context, means \emph{to get acquainted with}.
It is not to be confused with the \emph{familiarity thesis} that relates to the
projection of already known cognitive models onto the robot. We discuss this other
aspect at section~\ref{sec:cognitivemodel}.

The \emph{familiarization phase} starts at the peak of the \emph{novelty
effect}: anthropomorphic effects are at their maximum, the user thinks (s)he is
potentially facing an agent aiming at ``human-level intelligence'' (to take
McCarthy's words). This is a transient state that quickly vanishes, and the
projected anthropomorphism then starts to decrease while the human observes and
recognizes that the behaviour of the robot is generally \emph{predictable} and
possibly \emph{deceptive} (for instance, it talks but does not understand when
we talk ; it has eyes but does not recognize everyday objects we show ; etc.).

\paragraph{Disruptive behaviors}

By \emph{disruptive behaviors}, we mean any behavior exhibited by the robot
that is unexpected by the user: for instance, a robot may usually follow always
the same route to go to two different places in a house, and suddenly change
the route. The actual underlying reason may span from a bug to the detection of
a new possible obstacle, but as long as this reason is not immediately
intelligible to the user, the behavior counts as \emph{disruptive}.

Our model represents \emph{disruptive behaviors} as local increases of
anthropomorphic effects in the familiarization phase (and at a lesser extend,
in the stabilization phase) (Figure~\ref{fig:dynamics}): because such behaviors
are unexpected, a human observer interprets them as the result of a richer
deliberative process.

This is however to modulate depending on the real and percieved reasons of the
unexpected behavior.  Section~\ref{sec:disruptive} discusses more in depth the
mental correlates of disruptive behaviors. It proposes a typology that relates
them to the anthropomorphism they induce. It may increase if the user perceives
(rightfully or not) a form of intentionality, as well as decrease if the
unexpected behavior is perceived as a failure.
